---
title: "Group Separable BKMR Overview"
author: "by Danielle Demateis and Adon Osimo"
output: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{Group Separable BKMR Overview}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, echo = FALSE}
library(ggplot2)
```

# Introduction to bkmr and bkmrGS

The latest version of the standard bkmr package is on CRAN and can be
installed by

`install.packages("bkmr")`

The latest version of the bkmrGS package is on GitHub and can be
downloaded and installed by

`remotes::install_github("adon-osimo/bkmrGS")`

Then the packages can be loaded by

```{r}
library(bkmr)
library(bkmrGS)
```

This vignette introduces the `bkmr` and `bkmrGS` packages to implement
the standard and group-separable BKMR model. The `bkmr` package was
built by Jennifer Bobb and can be accessed
[here](https://CRAN.R-project.org/package=bkmr). The `bkmrGS` package is
built on top of the BKMR paper by Jennifer Bobb described in [this
Biostatistics article](https://doi.org/10.1093/biostatistics/kxu058) and
[this Environmental Health supplemental
article](https://doi.org/10.1186/s12940-018-0413-y).

Note that while this package is built off the `bkmr` package, it does
not have all of the functionality of that package. Currently the 
`'method = approx'` function argument has not been implemented for all functions
(see Changing methods in the User implemented changes section for more).  
If you are using both
the `bkmr` and `bkmrGS` packages in the same R session, make sure to
specify the package names when calling functions. For example, you can
call `bkmr::kmbayes` or `bkmrGS::bkmrGS` to differentiate
between the functions in the two packages, however you can run standard
BKMR through `bkmrGS::bkmrGS` by setting `mod = NULL`.

## Standard bkmr

Bayesian Kernel Machine Regression (BKMR) is a flexible framework for
modeling the health effects of exposure mixtures. It is designed to
estimate the joint effect of multiple exposures (Z) on an outcome (y),
while accounting for covariates (X). The standard BKMR model is

$$y_i = h(\mathbf{z_i}) + X_i^T\beta + \epsilon_i, \ \epsilon_i \sim N(0,\sigma^2)$$

where $h(\cdot)$ is an unknown function modeled using a kernel.  A common
choice is the Gaussian kernel with $m \in (1,\dots, M)$ exposure
variables,

$$K(\mathbf{z_i},\mathbf{z_j}, \tau, \mathbf{r}) = \tau\exp\left\{-\sum^M_{m=1}r_m(z_{im}-z_{jm})^2\right\}$$

Standard BKMR uses a global value of $\tau$, which controls the overall variability of the exposure-response function. It also uses a parameter $r$, which determines the contribution of each exposure to the kernel similarity. A larger value of $r_m$ implies a greater influence of exposure $m$ on the estimated exposure-response relationship. Intuitively, the kernel function shrinks the estimated health effects of
two individuals with similar exposure profiles toward each other [@bobb2015].

Standard BKMR assumes that this mixture-response function $h(\cdot)$ is
the same for all individuals, and does not account for effect
modification.

## Group Separable BKMR

The Group Separable BKMR framework extends standard BKMR by allowing us
to explicitly model effect modification in the kernel. While maintaining that each $i, j$ are individual observations, we add a categorical modifier $w_i \in (1,\dots, G)$, indicating which group the
observation belongs to. The model is

$$y_i = h_{w_i}(\mathbf{z_i}) + X_i^T\beta + \epsilon_i$$

Where each group $w_i$ has its own exposure-response function $h_{w_i}(\cdot)$. For every observation in group $w_i$, we have a new group Separable kernel in the form of

$$K_g(\mathbf{z_i, z_j}, w_i, \mathbf{\tau^{gs}, r^{gs}}) =\tau^{gs}_{w_i} \cdot \exp\Bigg[ - \sum_{m=1}^M {r^{gs}_{m,w_i}(z_{im} - z_{jm})^2} \Bigg]$$

The standard kernel and group separable kernel functions look very similar, still relying on the idea of shrinking similar observations within each exposure. However, in the group separable model, we have a different $\tau$, which we call $\tau^{gs}_{w_i}$ which varies under different groups. We also have a new $r$ which we call $r^{gs}_{m,w_i}$ which we also allow to vary by group. Under this formulation, observations only contribute to the kernel
within the same group, allowing each group to have a distinct and
flexible exposure-response surface.

[Does changing $K^{gs}$ to $K_g$ hinder the theme of both $\tau^{gs}, r^{gs}$? I know that this emphasizes the importance of the group being $g$]

## When to use Group Seperable BKMR and benefits of doing so

The classic BKMR model is appropriate when the exposure-response
relationship is assumed to be the same across all individuals. In
contrast, Groups Separable BKMR is useful when there is reason to
believe that this relationship differs across groups defined by a
categorical modifier $w$.

Compared to fitting separate BKMR models within each group (i.e.,
stratified analysis), Group Separable BKMR provides unified modeling.
This allows Group Separable BKMR to offer a more interpretable structure
by jointly estimating the group specific exposure-response functions
$h_w(Z)$. Group Separable BKMR is the right tool for fitting
exposure-response relationships across groups while maintaining
interpretability.

## Visualization of the Difference

```{r warning=FALSE, include=FALSE}
#make data for the plot
#ex_data_plot <- data.frame(y = Liu_data$y, Log_Lead = Liu_data$Log_Leag, Log_Manganese = Liu_data$Log_Manganese, Log_Arsenic = Liu_data$Log_Arsenic, ex_data$X)
set.seed(456)
data("Liu_data")
str(Liu_data)
y <- Liu_data$y
X <- Liu_data$Age
Z <- as.matrix(Liu_data[2:4])
#fit bkmr::kmbayes and get a plotable object
bkmr_model <- bkmr::kmbayes(y, Z, X, iter = 10, verbose = FALSE)
bkmr_input <- bkmr::PredictorResponseUnivar(fit = bkmr_model)
std_bkmr_plot <- ggplot(bkmr_input, aes(z, est, ymin = est - 1.96*se, ymax = est + 1.96*se)) + 
    geom_smooth(stat = "identity") + 
    facet_wrap(~ variable) +
  ylab("h(z)")

#fit bkmrGS::bkmrGS and get a plotable object
bkmrGS_model <- bkmrGS(y ~ h(Log_Lead, Log_Manganese, Log_Arsenic, mod = Sex) + Age, 
                          data = Liu_data,
                          iter = 10, 
                          verbose = FALSE) 
bkmrgs_input <- PredictorResponseUnivar(fit = bkmrGS_model, 
                                            method = "exact",
                                            q.fixed = 0.5)

bkmrGS_plot <- ggplot(bkmrgs_input, aes(z, est, ymin = est - 1.96*se, 
                                   ymax = est + 1.96*se, color = modifier, fill = modifier, linetype = modifier)) + 
  geom_smooth(stat = "identity") + 
  geom_hline(yintercept=0)+
  facet_wrap(vars(variable)) +
  xlab("Exposure level (log transformed, centered/scaled)") +
  ylab("Exposure-response function (h)") +
  theme_classic() +
  theme(legend.position = "bottom")+
  labs(color="Child Sex", fill = "Child Sex", linetype = "Child Sex") +
  scale_fill_manual(values=c("#E4A16C", "#789489"))+
  scale_color_manual(values=c("#D9782D", "#1E4D2B"))
```

Here is a quick demonstration of the difference between standard BKMR
and the group separable version of BKMR. In this example we know that
the exposure response relationship differs by child sex. However
standard BKMR ignores this relationship and assumes a common exposure
response surface across all subjects. This first plot uses our example
data (which we elaborate on later) and shows the results from improperly
fitting it with standard `bkmr`.

```{r, echo=FALSE}
std_bkmr_plot
```

Below is the result of fitting `bkmrGS` to that same data, allowing for
the exposure response function to vary by child sex group. Later in this
vignette, we discuss the code which generated these plots.

```{r, echo = FALSE}
bkmrGS_plot
```

The standard BKMR fit misses certain trends across child sex by
averaging across the two groups. Conversely, group separable BKMR is
able to fit the difference in the exposure response relationship for
males and females.

# Worked data example bkmrGS

In this section, we will provide a walk through of a data analysis
showcasing the usability of `bkmrGS`.

## Table of Contents

-   [Data Set and Fitting](#data-fitting)
-   [Model Fitting](#model-fitting)
-   [Model Checking](#model-checking)
-   [Model Output](#model-output)
-   [Exposure-Response Relationship](#exposure-response-relationships)
    -   [Univariate exposure-response functions](#univariate)
    -   [Group specific effects](#GSeffects)
    -   [Between Group Differences](#differences)

## Data Set and Fitting {#data-fitting}

Using the example data set in the package, we will fit the model using
the `bkmrGS()` function. First load the data set:

```{r}
data("Liu_data")
```

In this data from the @liu2021 paper we
have

-   `y` is the response (neurodevelopment score, BSID-III)

-   `Z` is a matrix of standardized exposures (Log_Lead, Log_Manganese,
    and Log_Arsenic measured in the fetus cumulatively during gestation)

-   `X` is a matrix of covariates (Age, Sex, Gestation, Delivery,
    Birth_order, Education_parent1, Education_parent2, Smoking,
    HOME_emotional, HOME_avoid, HOME_careg, HOME_env, HOME_play,
    HOME_stim, and Energy)

We used child sex (male/female) as a binary modifier of the effect of
exposure on neurodevelopment. Included is a small snippet of the data
frame

```{r}
head(Liu_data)
```

## Model Fitting {#model-fitting}

Now fit the model with the `bkmrGS()` function. The
following example uses 10 MCMC iterations, but it is important to use
ample MCMC iterations. Typically, we would like to see at least 1,000
usable iterations, but this largely depends on your effective sample
size and model convergence. To check model convergence, see [Model
Checking](#model-checking). Here, `bkmrGS()` takes the
arguments

-   `formula` a formula which takes the form of
    `y ~ h(Z_1,..., Z_k, mod = modifier_name) + X_1 +...+X_n` where

    -   `y`, the name of a column in your data frame that will be your
        response, note that it must be a column of a continuous variable

    -   `h(Z_1, Z_2, ... ,Z_k, mod = modifier_name)`, the names of every
        exposure you want included in the model (the `Z`'s), and the the name of
        the modifier column (`mod`). This is default set to NULL and runs standard BKMR.

    -   `X_1 + X_2 +...+ X_n`, the names of all covariates you would
        like added to the model separated by plus signs

-   `data`, the data frame which you are pulling all this from

-   `iter`, the number of iterations you want the MCMC sampler to run,
    which is defaulted to 1000.

-   `burnin`, the number of burn in iterations you want removed from
    your final sampling, which is defaulted to `iter/2`, which would be
    removing the first half of your samples

-   `verbose`, `TRUE` or `FALSE` (default) indicator for controlling the amount of
    live-time diagnostic which is outputted from the model,

-   `gs.tau`, `TRUE` or `FALSE` (default) indicator for whether to use
    group-specific $\tau$ parameters

-   `gs.sig`, `TRUE` (default) or `FALSE` indicator for whether to estimate
    separate error variance terms for each group

-   `est.h`, `TRUE` or `FALSE` (default) indicator for whether to sample from the
    posterior distribution of the subject-specific effects $h_i$ within
    the main sampler

More information about `gs.tau`, `gs.sig`, `est.h` please check their respective help
files. Other parameters (`id`, `starting.values`, `control.params`,
`varsel`, `groups`, `knots`, `ztest` and `rmethod`) are reused from the
standard `bkmr` package [@bkmr_overview]. Here we can see

```{r}
set.seed(111) #for consistency in demonstration
fitkm <- bkmrGS(y ~ h(Log_Lead, Log_Manganese, Log_Arsenic, mod = Sex) +
                          Age + Gestation + Delivery + Birth_order +
                          Education_parent1 + Education_parent2 + Smoking +
                          HOME_emotional + HOME_avoid + HOME_careg + 
                          HOME_env + HOME_play + HOME_stim + 
                          Energy, #entire model in a GAM style
                          data = Liu_data, #data frame
                          iter = 10, #MCMC iterations
                          verbose = FALSE) #controls output during fitting

```

We can quickly look at the object by printing it and using the
`summary()` function:

```{r}
summary(object = fitkm, #specify the model
        digits = 2 ) #specify the number of digits in output
```

In our example output, we caused several warnings to be tripped based
off our MCMC diagnostics. Acceptance rate warnings are displayed when
the MH acceptance rate is either below 20% or above 50%, as both
situations indicate poor mixing and possible convergence issues. We also
trip a warning for having our effective sample size is low (ESS \< 500),
very low (ESS \< 100), or unavailable. These all suggest potential
convergence problems. Finally there are two additional warnings related
to the number of retained posterior samples and burn-in. A warning will
be show if we have fewer than 1000 posterior draws, since small
posterior samples may lead to unstable estimates. The other warning is
indicated if the burn-in period is less than 10% of the total
iterations, meaning it might not have reached it's stationary
distribution yet. See [Model
Checking](#model-checking) for more details and trouble shooting.

## Model Checking {#model-checking}

Next, check the convergence of model parameters with the `TracePlot()`
function. This function asks you to specify the parameter with the `par`
argument and returns a line plot showing the sampled values of that parameter across MCMC iterations, along with the parameter's posterior mean displayed in orange. These trace plots should exhibits rapid up and down variation
with no long-term trends or drifts. This indicates that your estimates
have converged. Long-term trends or drifts in the plot indicate slower
convergence, poor convergence, or no convergence at all. See [Model
Checking](#model-checking) for more details and trouble shooting.

```{r, fig.width=8, fig.height=8}
par(mfrow = c(2, 2))
TracePlot(fitkm, par = "beta")
TracePlot(fitkm, par = "sigsq.eps")
TracePlot(fitkm, par = "lambda")
TracePlot(fitkm, par = "r", comp = 1)
```

Then, we should check the acceptance rates of $r$ and $\lambda$, to
ensure they are being sampled properly. In BKMR, the $r$ parameters
correspond to the exposures passed in the $h()$ call, and therefore
control the relative importance of each exposure's contribution. The
$\lambda$ parameter controls the overall flexibility of the exposure
response surface. In this code chunk, we are pulling the saved decision
(accept as 1 or reject as 0) from each iteration our model. Then we are
taking the column means which give the empirical acceptance rates for
each parameter

```{r}
colMeans(fitkm$acc.r)
colMeans(fitkm$acc.lambda)
```

Acceptance rates vary from problem to problem, but should not be close
to either 0 nor 1. A common rule of thumb is that the acceptance rates
for both $r$ and $\lambda$ should ideally be between 0.2 and 0.5.
Consider changing hyperparameters if acceptance rates are too high or
too low. See @bkmr_overview for more
details.

## Model Output {#model-output}

After reviewing our model to ensure proper mixing, and checking it's
acceptance rates, we can move on to reviewing the model output. Here if
we call the `ExtractEsts` function it will give us estimates for each of our
model parameters as well as quantiles for those estimates.

```{r}
ExtractEsts(fitkm, par = "beta", digits = 2)
ExtractEsts(fitkm, par = "sigsq.eps", digits = 2)
ExtractEsts(fitkm, par = "lambda", digits = 2)
ExtractEsts(fitkm, par = "r", digits = 2)
```

Similarly we
can use `ExtractSamps` to get and view all the samples our MCMC algorithm
decided to keep at each point

```{r}
samps <- ExtractSamps(fitkm, par = "beta")
head(samps)
```

Both functions take in a `par` argument, which indicates which parameter values to return. It is set `par = NULL` at default, and will return every parameter unless otherwise specified. 

## Exposure-Response Relationship {#exposure-response-relationships}

### Univariate exposure-response functions {#univariate}

I am still working on these sections, I guess I was a little confused as to what this reduction would look like. Would  I create
A new function that combines all the exposure-response relationships into one? Or just all the Overall vs Singular and Risk vs Interaction? 
Should it be one mega function? I guess when I left the meeting I thought I understood it but Im not as sure anymore. 
All of the functions in this section are run based off of ComputeMeanHnew just with different input parameters and over different values


To plot the estimated exposure-response function for each group, use the
`PredictorResponseUnivar()` function. This function takes a fit object
from `bkmrGS` (`fit`), estimation method (`"exact"` or
`"fullpost"`), and a reference quantile (`"q.fixed"`). Here `q.fixed` is
telling us the quantile value at which the exposures not being varied
are being held constant. This provides a reference point when estimating
the effect of changes in the exposure mixture. It returns the estimated
exposure-response function value (`est`) and standard error (`se`)
across a grid of exposure values (`z`) for each exposure (`variable`)
for each modifier level (`modifier`). This is particularly useful when
we want to visualize the relationship between our exposure and responses
for each individual group.

```{r}
pred.resp.univar <- PredictorResponseUnivar(fit = fitkm, 
                                            method = "exact",
                                            q.fixed = 0.5)
#this does return all three exposures, but I didn't want to use up a lot of lines in the pdf
#I understand that we want the user to see that this returns all of the exposure variables, but I couldn't think of a concice way to show that 
head(pred.resp.univar)
```

Then we can plot the output from `PredictorResponseUnivar()` to
visualize the differences in exposure-response functions across our
modifier. Each panel shows the estimated change in the outcome
associated with varying a single exposure while holding all other
exposures fixed at their median levels. Separate curves are shown for
each modifier group.

```{r}
ggplot(pred.resp.univar, aes(z, est, ymin = est - 1.96*se, 
                                   ymax = est + 1.96*se, color = modifier, fill = modifier, linetype = modifier)) + 
  geom_smooth(stat = "identity") + 
  geom_hline(yintercept=0)+
  facet_wrap(vars(variable)) +
  xlab("Exposure level (log transformed, centered/scaled)") +
  ylab("Exposure-response function (h)") +
  theme_classic() +
  theme(legend.position = "bottom")+
  labs(color="Child Sex", fill = "Child Sex", linetype = "Child Sex") +
  scale_fill_manual(values=c("#E4A16C", "#789489"))+
  scale_color_manual(values=c("#D9782D", "#1E4D2B"))
```

### Group specific effects {#GSeffects}

One of the primary advantages of the group-separable BKMR framework is
the ability to estimate group-specific mixture effects. Across each
level of the modifier, we can evaluate how the overall mixture and
individual exposures impact the outcome. In this section, we consider
two types of group-specific effects the overall mixture effect and the
single exposure effects. The former evaluates the change in outcome when
all exposures are increase simultaneously across qunatiles. The ladder
isolates the association between an individual exposure and the outcome
while holding the remaining exposures fixed. These summaries provide
insight into the joint behavior of the mixture and the individual
exposure contributions.

We begin by estimating the overall effect of the exposure mixture for a
fixed modifier level. The `OverallRiskSummaries()` function computes the
change in the outcome when all exposures are jointly increased to
specified quantiles, relative to a fixed reference level. This function
takes an argument `fit` which is an object from `bkmrGS`, the
argument `qs` specifies the quantiles at which the mixture is evaluated,
while `q.fixed` defines the reference quantile. The argument `m.fixed`
indicates the modifier level, so the resulting estimates correspond to
the group-specific function $h_w(Z)$

```{r}
total_mix_gs <- OverallRiskSummaries(fitkm,
                                     qs = c(0.25, 0.5, 0.75),
                                     q.fixed = 0.5,
                                     m.fixed = "male",
                                     method = "exact")
print(total_mix_gs)
```

One way to interpret this output that the `est` represent the estimated
change in outcome when the exposure mixture, for the fixed modifier
group, is jointly increased from the 50th quantile to the specified
quantile. If we isolate the 25th quantile for example, we would say that
among males, the estimated difference in the response when comparing the
exposure mixture at the 25th percentile versus the 50th percentile is
0.53 units. A positive estimate here indicates that the outcome is
higher at the flexible quantile compared to the fixed quantile. A
negative estimate would indicate the opposite, that the outcome is
higher at the fixed quantile compared to the flexible quantile.

To compare the overall mixture effects across groups, we repeat this
analysis for each modifier level and combine the results into a single
data frame for visualization.

```{r}
total_mix_gs$modifier <- rep("male", nrow(total_mix_gs))
total_mix_gs_tmp <- OverallRiskSummaries(fitkm,
                                         qs = c(0.25, 0.5, 0.75),
                                         q.fixed = 0.5,
                                         m.fixed = "female",
                                         sel = 5:10,
                                         method = "exact")
total_mix_gs_tmp$modifier <- rep("female", nrow(total_mix_gs_tmp))
total_mix_gs_df <- rbind(total_mix_gs, total_mix_gs_tmp)
```

The plot below shows how the estimated effect of increasing the entire
mixture varies by group. Differences between the curves indicate that
the overall mixture effect is not constant across levels of the
modifier, providing evidence of effect modification at the mixture
level.

```{r}
ggplot(total_mix_gs_df, 
       aes(x=quantile, y = est, ymin = est - 1.96*sd, 
           ymax = est + 1.96*sd, color = modifier, 
           shape = modifier))+
  geom_hline(yintercept=0)+
  geom_pointrange(position = position_dodge(width = 0.05), size=0.5)+
  theme_bw()+
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  xlab("Quantile")+
  ylab("Difference in BSID-III per \n IQR change in exposure") +
  labs(color="Child Sex", shape = "Child Sex") +
  scale_color_manual(values=c("#D9782D", "#1E4D2B"))

```

Next, we examine single exposure effects within the mixture using the
`SingVarRiskSummaries()` function. This function estimates the change in
the outcome associated with increasing one exposure from a lower to a
higher quantile, while holding the remaining exposures fixed. This
function takes an argument `fit` which is an object from
`bkmrGS`, the argument `qs.diff` specifies the lower and upper
quantiles used for this comparison, and `q.fixed` defines the quantile
at which the remaining exposures are held fixed. The argument `m.fixed`
indicates the modifier level, so the resulting estimates correspond to
the group specific exposure-response function $h_w(Z)$

```{r}

single_exp_gs <- SingVarRiskSummaries(fitkm,
                                       qs.diff = c(0.25, 0.75),
                                       q.fixed = 0.5,
                                       m.fixed = "male",
                                       method = "exact")
print(single_exp_gs)
```

The printed estimates here represent the estimated change in the outcome
associated with increase a single exposure. Specifically it represents
increasing the isolated exposure from the lower to upper quantile
specified while holding all other exposures fixed. Specifically in this
example, among males when increasing `Log_Lead` from the 25th percentile
to the 75th percentile is associated with a 0.262 unit change in the
outcome while holding the other exposures constant at their 50th
percentile.

To compare effects across groups, we repeat this analysis for each
modifier level and combine the results into a single data frame for
visualization.

```{r}
single_exp_gs$modifier <- rep("male",nrow(single_exp_gs))

single_exp_gs_tmp <- SingVarRiskSummaries(fitkm,
                                            qs.diff = c(0.25, 0.75),
                                            q.fixed = 0.5,
                                            m.fixed = "female",
                                            sel = 5:10,
                                            method = "exact")
single_exp_gs_tmp$modifier <- rep("female",nrow(single_exp_gs_tmp))
single_exp_gs_df <- rbind(single_exp_gs, single_exp_gs_tmp)
```

The following plot displays the estimated change in the outcome for an
interquartile range (IQR) increase in each exposure, stratified by
modifier group. Differences between groups reflect effect modification
in the exposure-response relationship.

```{r}
ggplot(single_exp_gs_df, 
       aes(x=q.fixed, y = est, ymin = est - 1.96*sd, 
           ymax = est + 1.96*sd, color = modifier, 
           shape = modifier))+
  geom_hline(yintercept=0)+
  geom_pointrange(position = position_dodge(width = 0.5), size=0.5)+
  facet_wrap(vars(variable))+
  theme_bw()+
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  xlab("Quantile")+
  ylab("Difference in BSID-III per \n IQR change in exposure") +
  labs(color="Child sex", shape = "Child sex") +
  scale_color_manual(values=c("#D9782D", "#1E4D2B"))

```

### Between Group Differences {#differences}

In addition to estimating group-specific effects, Group-Separable BKMR
allows for direct comparison of exposure effects between groups. This
provides a formal way to assess effect modification, by quantifying how
the exposure-response relationship differs across the modifier.

We estimate the difference in the overall mixture effect between groups
using the `OverallIntSummaries()` function. This function computes the
difference in the estimated mixture effect between two modifier levels
at specified quantiles of the exposure distribution. This function takes
an argument `fit` which is an object from `bkmrGS`, the argument
`mod.diff = c("male", "female")` indicates that the estimated
differences correspond to female minus male, the argument `sel = 5:10`
specifies which posterior samples from the fitted model are used when
computing the summaries. Positive values therefore indicate a stronger
association in females, while negative values indicate a stronger
association in males. The argument `qs` specifies the quantiles at which
the mixture is evaluated, while `q.fixed` defines the reference
quantile.

```{r}
total_mix_btwn <- OverallIntSummaries(fitkm,
                                             qs = c(0.25, 0.5, 0.75),
                                             q.fixed = 0.5,
                                             mod.diff = c("male", "female"), #female - male
                                             sel = 5:10,
                                             method = "exact")
print(total_mix_btwn)
```

This output could be interpreted as the estimated difference in the
overall mixture effect between females and males when simultaneously
changing all exposures from the reference quantile to the specified
quantile. For example, we could say that for the 75th percentile, that
the estimated effect of increase the overall exposure mixture is 0.517
units larger in females than in males at that quantile

The following plot displays the difference in overall mixture effects
between groups across quantiles. Confidence intervals that exclude zero
provide evidence of effect modification at the mixture level.

```{r}
ggplot(total_mix_btwn, aes(x=quantile, y = est, ymin = est - 1.96*sd, 
                                   ymax = est + 1.96*sd))+
  geom_hline(yintercept=0)+
  geom_pointrange(size=0.5)+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  xlab("Quantile")+
  ylab("Difference in association \n (per IQR change) with BSID-III \n between girls and boys")
```

To estimate differences in single-exposure effects between groups, we
use the `SingVarIntSummaries()` function. This function evaluates how
the effect of changing one exposure differs across modifier levels,
while holding all other exposures fixed. This function takes an argument
`fit` which is an object from `bkmrGS`, the argument `qs.diff` a
vector of two quantiles for computing the difference `qs.diff`, the
argument `q.fixed` which is a single value repeated twice in a vector at
which to fix the other exposure, and `m.fixed` which is the modifier
level. The following example computes the estimated difference in effect
between boys and girls of an IQR change in each exposure when the others
are fixed at the median.

```{r}
single_exp_btwn <- SingVarIntSummaries(fitkm,
                                       qs.diff = c(0.25, 0.75),
                                       qs.fixed = c(0.5,0.5),
                                       mod.diff = c("male", "female"), #female - male
                                       method = "exact")
print(single_exp_btwn)
```

In this example, the estimate for `Log_Lead` is -0.406, which suggests
that the estimated effect of increase `Log_Lead` from the 25th
percentile to the 75th percentile is 0.406 units smaller in females
compared to males while holding the other exposures fixed at their 50th
percentiles.

Similarly, we can visualize differences in single-exposure effects
between groups. Each panel corresponds to an exposure, and the plotted
estimates represent the difference in effect (female minus male) for an
IQR increase in that exposure.

```{r}
ggplot(single_exp_btwn, aes(x="", y = est, ymin = est - 1.96*sd, 
                                   ymax = est + 1.96*sd))+
  geom_hline(yintercept=0)+
  geom_pointrange(size=0.5)+
  facet_wrap(vars(variable))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  xlab("")+
  ylab("Difference in association \n (per IQR change) with BSID-III \n between girls and boys")
```

# User Implemented Changes

## How to Change Estimation Methods

Several functions in `bkmrGS` include a `method` argument that controls
how risk summaries are computed. Different methods may vary in
computational cost and the degree of approximation used. Specifically,
all of our exposure-response functions have the `method` argument, which
can be `approx`, `exact`, or `fullpost`. The default for this is
changes depending on the function (check the help fuile for each function if 
you need the default). When `method = approx`, the function uses a 
Gaussian process approximation based on
posterior mean and variance estimates. This reduces the computational
burden slightly and scales better when there are many prediction points. If you
changed this to `method = exact` we see that the function is evaluated
using a full Gaussian process. It also uses the exact covariance
structure at the posterior mean parameter estimates. This is more
precise at the cost of being more computationally expensive. Between
`exact` and `approx` the only difference is that `approx` uses an
approximation of the Gaussian process and the `exact` uses the actual
Gaussian process. Finally the `method = fullpost` is only available for
`PredictorResponseUnivar`. This computes the predictions separately at
each MCMC iteration, and then summarizes these predictions across
posterior samples.

The function `bkmrGS()` also has a `kernel.method` call,
which allows the user to run either standard BKMR if
`kernel.method = "one"` or group separable BKMR if
`kernel.method = "two"`. Note that if you include a modifier you should be running  `kernel.method = "two"`, and that if you don't include a modifier `kernel.method = "one"` is set automatically. 

For example, in functions such as `OverallRiskSummaries()` and
`SingVarRiskSummaries()`, the argument `method = "exact"` computes risk
differences directly using posterior samples of the exposure-response
function. This approach is more accurate but can be computationally
intensive. Alternative methods (e.g., `method = "approx"`) may use
approximations to reduce computation time. These approaches can be
useful for large data sets or when rapid exploratory analysis is
desired, but may introduce additional variability or bias.

## How to Change the Tuning Parameters

The BKMR model is fit using Markov Chain Monte Carlo, therefore most of the parameters are updated using a Gibbs step. The $r_m$ and $\lambda$ parameters, however are updated using a Metropolis-Hastings step. We used a gamma proposal distribution with mean equal to the current value. 

If your acceptance rates are too high (an acceptance rate higher than 50\%) this could lead to a slow mixing chain. If your acceptance rates are too low (an acceptance rate lower than 20\%), that could prevent the algorithm from properly exploring the parameter space. Thus it is essential for any MCMC to have good acceptance rates. Generally, these acceptance rate can be changed through the `control.params` argument. If you are running standard BKMR, then you will only see a list of 1 value per parameter, but if you are running group-seperable bkmr, then you can change the tuning parameter per group. Here is the list of tuning parameters and thier starting values

```{r}
data.frame(fitkm$control.params)
```

Any value that ends in a `.jump` or `.muprop` is a tuning parameter and can be changed. For more information see @bkmr_overview

## How to Change the Priors

Our package also allows users to modify prior distributions for the model
parameters. Prior specification can be updated again through the
`control.params` argument. Priors can be explored using the
`InvestigatePriors()` function, which allows users to visualize the
implied distributions under different hyperparameter choices. After
specifying alternative priors, the `PlotPriorFits()` function can be
used to visualize how these choices influence the implied
exposure-response relationships. Here is the list of prior hyperparameters and their starting values

```{r}
data.frame(fitkm$control.params)
```

Any value that does not end in a `.jump` or `.muprop` is a prior parameter and can be changed. For more information see @bkmr_overview

## How to Handle Poor Mixing Results

Model convergence is essential when dealing with any form of Monte Carlo Markov Chain estimation. This allows us to be confident in our estimates, as our algorithm has stabilized around an optimal state. A few ways to test model convergence are through the `TracePlot` function and the effective sample size (`ess`) return when calling summary on a `bkmrGS` object. 

To check the effectiveness of your sampling, a trace plot is ideal. Like we talked about in the Model Checking Section, we want to see no obvious trends (post burn in) and a rapid up and down variation with no drifts. If this is not the case, either increase the number number of samples (try `iter = 1500,...,burnin = 500` in the `bkmrGS()` call or larger) or adjust the tuning parameters. 

If your effective sample size is too low, this indicates your chains are producing highly correlated estimates, and it is taking a long time to explore your parameter space. You would know if your effective sample size is too low if the summary produces a warning. If this is the case, either try increasing the number of samples (try `iter = 1500, ... , burnin = 500` in the `bkmrGS()` call or larger), adjusting the tuning parameters, or thinning. Thinning is the process of saving every $n$-th iteration instead of saving every iteration. Typically, increasing iterations or improving mixing is generally preferable to thinning.

[I should probably add a call for this somewhere in the function, like `thin = _`]

# References










